SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-12-04 design_arch.tex
Completing Methodology, Images

In addition to the prefetch table mentioned above, PPF also maintains
a ``Reject Table.''  The reject table is a 1024-entry deep
direct-mapped table.  If a prefetch suggestion is rejected by the
perceptron layer, it is logged into the reject table.  The filer is
used to train the perceptron to %Gino: Filer?
avoid false negatives \textit{i.e.}, cases where prediction was to
reject the prefetch but the prefetch would have been useful.

>>>>

In addition to the prefetch table mentioned above, PPF also maintains
a ``Reject Table.''  The reject table is a 1024-entry deep
direct-mapped table.  If a prefetch suggestion is rejected by the
perceptron layer, it is logged into the reject table.  The table is
used to train the perceptron to %Gino: Filer? [EB]: Resolved
avoid false negatives \textit{i.e.}, cases where prediction was to
reject the prefetch but the prefetch would have been useful.

### 2018-12-04 design_impl.tex
Completing Methodology, Images

%%[NOTE] Can this point be removed?
% djimenez: I think this would be necessary to reproduce our results so leave it in.

\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
  original SPP does not consider the slots available in the L2 MSHR
  queue.  It has the assumption that at no time can the number of
  suggested prefetches exceed the capacity of the queue.  While it was
  not described in the paper, the internal confidence mechanism makes
  sure that the above assumption was maintained.  In PPF, since there
  cannot be an assumption on the degree of aggressiveness of the
  underlying prefetcher, we explicitly check that at no point should
  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
    this a big deal?  Seems like a detail}

\item \textbf{Enhanced Prefetch Table}\newline The prefetch table
  introduced in the original SPP keeps track of up to 1024 L2 cache
  prefetch suggestions.  For PPF, the prefetch table is modified to
  store metadata such as program counter, memory access address etc.
  This metadata is needed to retrieve the
  state %Gino: What's the etc.?
  of the features to train the perceptron at a later stage in
  time. The prefetch table mentioned above keeps track of the
  prefetches that actually happened in the L2 Cache.  We introduce the
  reject table to keep a track of the prefetches that were suggested
  by the base SPP engine but were rejected by PPF.  In case it is
  detected that a future demand fetch could have used this rejected
  prefetch, the perceptron weights are updated to reflect that
  misprediction.

>>>>

% EB: Can this point be removed?
% djimenez: I think this would be necessary to reproduce our results so leave it in. 
% EB: It's ultimately a check to avoid SPP's internal segfault. Won't have a real impact.

%\item \textbf{Looking at L2 MSHR Queue while prefetching}\newline The
%  original SPP does not consider the slots available in the L2 MSHR
%  queue.  It has the assumption that at no time can the number of
%  suggested prefetches exceed the capacity of the queue.  While it was
%  not described in the paper, the internal confidence mechanism makes
%  sure that the above assumption was maintained.  In PPF, since there
%  cannot be an assumption on the degree of aggressiveness of the
%  underlying prefetcher, we explicitly check that at no point should
%  the number of suggested prefetches exceed the L2 MSHR queue. \pg{is
%    this a big deal?  Seems like a detail}

\item \textbf{Enhanced Prefetch Table}\newline The prefetch table
  introduced in the original SPP keeps track of up to 1024 L2 cache
  prefetch suggestions.  For PPF, the prefetch table is modified to
  store the features used for perceptron indexing..
  This metadata is needed to retrieve the
  values of the features to train the perceptron at a later stage in
  time. The prefetch table mentioned above keeps track of the
  prefetches that actually happened in the L2 Cache.  We introduce the
  reject table to keep a track of the prefetches that were suggested
  by the base SPP engine but were rejected by PPF.  In case it is
  detected that a future demand fetch could have used this rejected
  prefetch, the perceptron weights are updated to reflect that
  misprediction.
%Gino: What's the etc.?
%EB: Resolved
\end{itemize}

### 2018-12-04 method.tex
Results Section and PDF Images
Completing Methodology, Images

\begin{table}[]
    \centering
    \begin{tabular}{|l|p{3.6cm}|c|}
    \hline
    Level & Configuration & Access Latency \\
    \hline
         L1 Cache & Separate I-cache and D-cache, 32 KB, 8-way & 4 cycles\\
         L2 Cache & Private, 256 KB & 8 cycles\\
         LLC & Shared, 2MB / core, 16-way & 20 cycles\\
         DRAM & 4 GB Single Channel for single-core, 8 GB Double Channel for multi-core & N?? cycles\\
    \hline
    \end{tabular}
    \caption{Memory Model Parameters}
    \label{tab:Sim_params}
\end{table}

@@

The block size is fixed at 64 bytes. Prefetching is only triggered at an L2 cache
demand accesses but could be directed to the L2 or last-level cache. No L1
data level prefetching is done. The LRU replacement policy is used on all levels
of cache hierarchies. Branch prediction is done using the perceptron branch
predictor~\cite{Perc_Branch}.  The page size is 4KB.  Champsim operates
all the prefetchers strictly in the physical address space.

@@

In total, we used 285 traces representing workloads across 53 applications.
Throughout the paper for SPEC 2017, we consider memory intensive subset as the
traces with LLC MPKI > \_\_.  This includes 48 out of the 95 simpoints
developed.  Since SPEC 2006 differs considerably in terms of memory behaviour,
we define memory intensive subset as traces with LLC MPKI > \_\_.  This
includes 47 out of 94 simpoints.

@@

We plot a histogram for each feature depicting weights distribution
from -16 to +15 and use this histogram to generate the Pearson's
correlation factor for that feature.  In statistics, Pearson's factor
is a numerical measure of the degree of linear correlation between two
variables.  It ranges from -1 to +1.  The magnitude of Pearson's
factor depicts the extent of correlation and the sign depicts whether
it is a positive correlation or a negative correlation.  Values close
to 0 suggest a low correlation or at times, noisy data.  A value of
+1/-1 suggests a perfectly linear positive / negative correlation
respectively.  Figure XX depicts the histogram and the Pearson's
coefficient for two of the features, YY and ZZ.

Figure XX shows all the features used, arranged in the increasing
order of their Pearson's factor.  As can be seen X out of the 9
features provide a moderate to high correlation, with the magnitude of
P-value > 0.4.  The single most important feature, \_\_ helps provide
a correlation to prefetch outcome with a factor of \_\_.

\textbf{Per Trace Correlation} \newline Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces.  Here we give special attention to features
with low P-values in the previous experiment.  Figure XX shows the
variation of P-values three features : \_\_, \_\_ and \_\_; across all
the SPEC 2017 traces.  For simplicity, the traces are arranged in an
increasing order of contribution made by the feature and the trace
names have been omitted.  It can be seen that even features with a low
overall correlation provide useful correlation (magnitude > 0.4) for
XX out of the 95 simpoints developed for SPEC 2017.

@@

\textit{<TODO: VALIDATE THIS>} Secondly, studying the relative
importance of each feature enabled us to vary the number of entries
dedicated for each feature.  Features with higher Pearson's
correlation (like \_\_ and \_\_) were given most importance and
allowed full 12-bits of indexing.  Features like \_\_ and \_\_, with a
low overall P-value were allocated fewer indexes in the feature table.

@@

\subsection{Overhead for PPF}
\label{Method-Overheads}

\begin{table}[t]
    \centering
    \begin{tabular}{|c|c|m{4.8cm}|}
    \hline
        \textbf{Field} &
        \textbf{Bits} &
        \textbf{Comment} \\
    \hline
         Valid & 1 & Indicates a valid entry in the table\\
         Tag & 6 & Identifier for the entry in the table\\
         Useful & 1 & To show if the given entry led to a useful demand fetch\\
         Perc Decision & 1 & Prefetched vs Not-prefetched \\
    \hline
        PC & 12 & \multirow{5}{4.8cm}{Meta-data required for perceptron training}\\
        Address & 24 & \\
        Delta & 7 & \\
        Curr Signature & 12 & \\
        Confidence & 7 & \\
    \hline
        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
    \hline
    \end{tabular}
    \caption{Meta-data Stored in Prefetch Table}
    \label{tab:PTable_metadata}
\end{table}

@@

\begin{table}[h]
\begin{adjustwidth}{-0.5cm}{}
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
        \textbf{Structure} &
        \textbf{Entry} &
        \textbf{Components} &
        \textbf{Total} \\
    \hline
        \multirow{5}{1.5cm}{Signature Table} &      & Valid (1 bit)        &             \\
                                             &      & Tag (16 bits)        &             \\
                                             & 256  & Last Offset (6 bits) & 11008 bits  \\  
                                             &      & Signature (12 bits)  &             \\
                                             &      & LRU (6 bits)         &             \\
    \hline
        \multirow{3}{2cm}{~~Pattern Table} &      & $C_{sig}$ (4bits)      &               \\
                                             & 512  & $C_{delta}$ (4*4 bits) & 24576 bits    \\
                                             &      & Delta (4*7 bits)       &               \\
    \hline
        \multirow{3}{1.5cm}{Perceptron\newline}     & 4096*8    &            &              \\
        \multirow{2}{0.9cm}{Weights}                 & 2048*x    & 5 bits     & 164480 bits  \\
                                                    & 128*1     &            &              \\
    \hline
        Prefetch Table\footnotemark[1]             & 1024      & 84 bits    & 86016 bits   \\
    \hline
        Reject Table\footnotemark[2]               & 1024      & 83 bits    & 84992 bits   \\
    \hline
        \multirow{4}{1cm}{Global\newline\newline}   & \multirow{4}{0.2cm}{8} & Signature (12 bits)  & \multirow{4}{1.1cm}{264 bits} \\
        \multirow{3}{1.1cm}{History\newline}        &                        & Confidence (8 bits)  &                               \\
        \multirow{2}{1.2cm}{Register}               &                        & Last Offset (6 bits) &                               \\
                                                    &                        & Delta (7 bits)       &                               \\
    \hline
        Accuracy        & 1     & C$_{total}$       & 10 bits   \\
        Counters        & 1     & C$_{useful}$      & 10 bits   \\
    \hline
        \multirow{3}{1.5cm}{Global PC\newline}      &       & $PC_1$ (12 bits)      &           \\
        \multirow{2}{1.5cm}{~Trackers}              & 3     & $PC_2$ (12 bits)      & 36 bits   \\
                                                    &       & $PC_3$ (12 bits)      &           \\
    \hline
        \multicolumn{4}{|c|}{\textbf{Total: 371392 bits = 45.33 KB}}\\
    \hline
    \end{tabular}
    \caption{SPP-Perc Storage Overhead}
    \label{tab:PPF_overhead}
\end{adjustwidth}
\end{table}

@@

\begin{figure}[h]
\includegraphics[width=\columnwidth]{SPEC2017}
\caption{SPEC 2017 Single-Core IPC Speedup}
\label{Fig:SPEC2017_1core}
\end{figure}

>>>>

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|}
    \hline
    	 Block & Configuration\\
    \hline
	 \multirow{2}{1.5cm}{CPU Core} 		& 1-8 Cores, 4 GHz\\
						& 256 entry ROB, 4-wide\\
    \hline
         \multirow{2}{2.7cm}{Private L1 DCache} & 32 KB, 8-way, 4 cycles\\
						& 8 MSHRs, LRU\\
    \hline
         \multirow{2}{2.4cm}{Private L2 Cache}  & 256 KB, 8-way, 8 cycles\\
						& 16 MSHRs, LRU, Non-inclusive\\		
    \hline
         \multirow{2}{1.7cm}{Shared LLC} 	& 2MB/core, 16-way, 12 cycles\\
						& 32 MSHRs, LRU, Non-inclusive\\
    \hline
         \multirow{3}{1.1cm}{DRAM} 		& 4 GB 1-Channel (single-core)\\ 
						& 8 GB 2-Channels (multi-core)\\ 
						& 64-bit channel, 1600MT/s\\
    \hline
    \end{tabular}
    \caption{Simulation Parameters}
    \label{tab:Sim_params}
\end{table}

@@

The block size is fixed at 64 bytes. Prefetching is only triggered at an L2 cache
demand accesses but could be directed to the L2 or last-level cache. No L1
data level prefetching is done. The LRU replacement policy is used on all levels
of cache hierarchies. Branch prediction is done using the perceptron branch
predictor~\cite{Perc_Branch}.  The page size is 4KB.  ChampSim operates
all the prefetchers strictly in the physical address space.

@@

In total, we used 285 traces representing workloads across 53 applications.
Throughout the paper, we consider memory intensive subset as the
traces with LLC MPKI > 1.  This includes 48 out of the 95 simpoints
developed for SPEC 2017. For SPEC 2006  This includes 48 out of 94 simpoints.

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\columnwidth]{Hist}
    \caption{Distribution of Trained Weights}
    \label{Fig:Hist}
  \end{center}
\end{figure}

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\columnwidth]{Pvals}
    \caption{P-Values for all Features}
    \label{Fig:Pvals}
  \end{center}
\end{figure}

\begin{figure}[h]
  \begin{center}
    \includegraphics[width=\columnwidth]{Pvals_Trace}
    \caption{P-value Variation across Traces for selected Features}
    \label{Fig:Pvals_Trace}
  \end{center}
\end{figure}

We plot a histogram for each feature depicting weights distribution
from -16 to +15 and use this histogram to generate the Pearson's
correlation factor for that feature.  In statistics, Pearson's factor
is a numerical measure of the degree of linear correlation between two
variables.  It ranges from -1 to +1.  The magnitude of Pearson's
factor depicts the extent of correlation and the sign depicts whether
it is a positive correlation or a negative correlation.  Values close
to 0 suggest a low correlation or at times, noisy data.  A value of
+1/-1 suggests a perfectly linear positive / negative correlation
respectively.  Figure~\ref{Fig:Hist} depicts the histogram distribution 
of trained weights, for a feature with high and another with moderate
Pearson's coefficient.

Figure~\ref{Fig:Pvals} shows all the features used, arranged in the increasing
order of their Pearson's factor.  As can be seen 5 out of the 9
features provide a moderate to high correlation, with the magnitude of
P-value > 0.6.  The single most important feature, 
\textit{Confidence XOR Page address} helps provide
a correlation to prefetch outcome with a factor of 0.90.

\textbf{Per Trace Correlation} \newline Another important way to look
at the perceptron features is to see how much their contribution
varies across the traces.  Here we give special attention to features
with low P-values in.  Figure~\ref{Pvals_Traces} shows the
variation of P-values for three features : \textit{PC XOR Delta},
\textit{Signature XOR Delta} and 
\textit{PC XOR Depth}; across all the SPEC 2017 traces.  
For simplicity, the traces are arranged in an
increasing order of contribution made by the feature and the trace
names have been omitted.  It can be seen that even features with a low
overall correlation provide useful correlation (magnitude > 0.5) for
a significant number of traces.

@@

Secondly, studying the relative importance of each feature enabled 
us to vary the number of entries dedicated for each feature.  
Features with higher Pearson's correlation, \textit{cache line} 
and \textit{page address}  were given most importance and
allowed full 12-bits of indexing.  Features like \textit{ip XOR delta} 
and \textit{ip XOR depth} with a low overall P-value 
were allocated fewer indexes in the feature table.

@@

\begin{table}[ht]
    \centering
    \begin{tabular}{|c|c|c|}
    \hline
        \textbf{Field} &
        \textbf{Bits} &
        \textbf{Comment} \\
    \hline
         Valid 		& 1  & Indicates a valid entry in the table\\
         Tag 		& 6  & Identifier for the entry in the table\\
         \multirow{2}{1cm}{Useful} 	& 1  & To show if the given entry led to a\\
                    	&    & useful demand fetch\\
         Perc Decision 	& 1  & Prefetched vs Not-prefetched \\
    \hline
        PC 		& 12 & \multirow{4}{4.8cm}{Meta-data required for perceptron}\\
        Address 	& 24 & \multirow{4}{1.1cm}{training}\\
        Delta 		& 7  & \\
        Curr Signature 	& 12 & \\
        Confidence 	& 7  & \\
    \hline
        \multicolumn{3}{|c|}{\textbf{Total 84 bits}}\\
    \hline
    \end{tabular}
    \caption{Meta-data Stored in Prefetch Table}
    \label{tab:PTable_metadata}
\end{table}

@@

\begin{table}[h]
\begin{adjustwidth}{-0.1cm}{}
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
        \textbf{Structure} &
        \textbf{Entry} &
        \textbf{Components} &
        \textbf{Total} \\
    \hline
                                            &  \multirow{5}{0.5cm}{256}    & Valid (1 bit)  &             \\
                                             &      & Tag (16 bits)        &  \multirow{2}{0.9cm}{11008}           \\
                            Signature Table  &   & Last Offset (6 bits) &  \multirow{2}{0.5cm}{bits}  \\  
                                             &      & Signature (12 bits)  &             \\
                                             &      & LRU (6 bits)         &             \\
    \hline
                                    &  \multirow{3}{0.5cm}{512}    & $C_{sig}$ (4bits)      &\multirow{2}{0.9cm}{24576}               \\
                       Pattern Table         &   & $C_{delta}$ (4*4 bits) &  \multirow{2}{0.5cm}{bits}  \\
                                             &      & Delta (4*7 bits)       &               \\
    \hline
        \multirow{4}{1.5cm}{Perceptron\newline}     & 4096*4    & \multirow{4}{0.8cm}{5 bits}  & \multirow{3}{1.1cm}{113280}             \\
        \multirow{3}{1.2cm}{Weights}                & 2048*2    &           &  \multirow{3}{0.5cm}{bits}  \\
                                                    & 1024*2    &           &               \\
                                                    & 128*1     &           &              \\
    \hline
        Prefetch                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{85 bits}       & 87040 \\
        Table\footnotemark[1]   &           &               & bits\\
    \hline
        Reject                & \multirow{2}{0.7cm}{1024}      & \multirow{2}{1cm}{84 bits}    & 86016 \\
        Table\footnotemark[2] & & & bits\\
    \hline
        \multirow{4}{1.0cm}{Global\newline\newline}   & \multirow{4}{0.2cm}{8} & Signature (12 bits)  & \multirow{4}{1.1cm}{264 bits} \\
        \multirow{3}{1.1cm}{History\newline}        &                        & Confidence (8 bits)  &                               \\
        \multirow{2}{1.2cm}{Register}               &                        & Last Offset (6 bits) &                               \\
                                                    &                        & Delta (7 bits)       &                               \\
    \hline
        Accuracy        & 1     & C$_{total}$       & 10 bits   \\
        Counters        & 1     & C$_{useful}$      & 10 bits   \\
    \hline
        \multirow{3}{1.5cm}{Global PC\newline}      &       & $PC_1$ (12 bits)      &           \\
        \multirow{2}{1.5cm}{~Trackers}              & 3     & $PC_2$ (12 bits)      & 36 bits   \\
                                                    &       & $PC_3$ (12 bits)      &           \\
    \hline
        \multicolumn{4}{|c|}{\textbf{Total: 3222940 bits = 39.34 KB}}\\
    \hline
    \end{tabular}
    \caption{SPP-Perc Storage Overhead}
    \label{tab:PPF_overhead}
\end{adjustwidth}
\end{table}

@@
